### 2015-10-31 models/priors-consortium.Rmd
Print posterior over parameters
Posterior predictives still need some thought.
How do you get out a posterior predictive for the dirichlet?

The posterior over parameters examined using 50k samples are only vaguely reasonable.
The independence between states (dirichlet prior) may not be valid.

Also, still need to analyze the human data to be able to compare it to the posterior predictive.

---
title: "priors-consortium"
author: "mht"
date: "October 31, 2015"
output: html_document
---

```{r helpers}

library(coda)
library(hydroGOF)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}
HPDhi<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}
HPDlo<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
options("scipen"=10)   
```

# Human data

```{r data.load}
d.number <- read.csv("~/Documents/research/priors/prag/data/number_dat.csv")
d.binned <- read.csv("~/Documents/research/priors/prag/data/bin_dat_noZeros.csv")
d.choice <- read.csv("~/Documents/research/priors/prag/data/choice_dat.csv")
```

```{r data.number}
qplot(data=d.number, x = chosen_bin, geom='histogram')+
  facet_wrap(~tag)

```

```{r data.binned}
d.bin.tidy<- d.binned %>%
  group_by(tag, bin_num) %>%
  multi_boot_standard(column="nresponse")

ggplot(d.bin.tidy, aes(x=bin_num, y = mean))+
  geom_bar(stat='identity', position=position_dodge())+
  geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper))+
  facet_wrap(~tag)
```

```{r data.choice}

```

# Model

```{r model.load}
fpath<- "~/Documents/research/priors/models/results/"
prefix<- "priors-base-IncrMH"
samples = 50000
burn = samples /2
m<-read.csv(paste(fpath, prefix, samples, "burn", burn,".csv", sep=''))
m.samples <- m[rep(row.names(m), m$Probability*samples), 1:18]
```

# Posterior over parameters

## Data analysis parameters

```{r posterior.parameters}

m.params <- m.samples %>%
  filter(Type=='parameter')

ggplot(m.params, aes(x=Value))+
  geom_histogram()+
  facet_wrap(~Parameter, scales='free')

```

## Parameters governing the big prior in the sky

```{r}
m.pp <- m.samples %>%
  filter(Type == 'posterior' & Parameter=='priorVector') %>%
  gather(Bin, Value, starts_with("X")) %>%
  group_by(Parameter, Item, Bin) %>%
  summarize(MAP = estimate_mode(Value),
            credHigh = HPDhi(Value),
            credLow = HPDlo(Value)) %>%
  mutate(Bin = as.numeric(as.character(gsub("X","",Bin))))

ggplot(m.pp, aes(x=Bin, y=MAP))+
  geom_bar(stat='identity', position=position_dodge())+
  geom_errorbar(aes(ymin=credLow, ymax=credHigh))+
  facet_wrap(~Item)

```